### 2016-11-23 thesis/intro/intro.tex
tmp commit

@@

\subsubsection{Metastasis}
\label{ssub:metastasis}

\subsubsection{Tumour heterogeneity}
\label{subsub:tumour_heterogeneity}

>>>>

\noindent
Many of the hallmarks are acquired by the cancer cells during its growth, through many genomic changes and mutations of the essential genes that allow cancer cells to survive, proliferate, invade and metastasise.
Thus, it is only logical to induce that many, if not all, cancers are more prone to genetic mutations and restructuring, which allows the cancers to subsequently acquire the already described Hallmarks of Cancer \citep{Hanahan2011}.

Until recently, the evolutionary progress of cancers have been thought to be gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
There are now evidence that show otherwise, whereby a single catastrophic event in the chromosome structure occurs, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
This rare event that is estimated to occur in 2$\sim$5\% of all cancers was termed \gls{chromothripsis}, and is thought to be caused by the physical separation of parts of chromosome by a structure called micronuclei \citep{Leibowitz2015,Stephens2011}.
Owing to its nature of being a one-off chromosomal disruption, chromothripsis (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.

In addition to this, areas of the chromosome where major structural rearrangements have occured can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
By deaminating cytidine to uracil in the \gls{ssdna} that may occur at the sites of double strand break, \gls{aid}/\acrshort{apobec} family proteins introduces hypermutated regions in the genome \citep{Leibowitz2015,Nik-Zainal2012}.

It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately contribute to many of the tumorigenic phenotypes.

@@

% TODO: do I really need tumour heterogeneity??
\subsubsection{Tumour heterogeneity}
\label{subsub:tumour_heterogeneity}

### 2016-11-23 thesis/misc/glossary.tex
tmp commit

\newglossaryentry{ldp}{
	name=Limited duration prevalence,
	text=limited duration prevalence,
	description={Number of patients diagnosed with a disease within a fixed period of time in the past}
}

@@

@@

\newacronym{blca}{BLCA}{bladder urothelial carcinoma}
\newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
\newacronym{coad}{COAD}{colon adenocarcinoma}
\newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
\newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
\newacronym{read}{READ}{rectum adenocarcinoma}
\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
\newacronym{geo}{GEO}{Gene Expression Omnibus}
\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
\newacronym{bmi}{BMI}{Body Mass Index}
\newacronym{who}{WHO}{World Health Organization}
\newacronym{ln}{LN}{Lymph node}
\newacronym{pr}{PR}{Progesterone receptor}
\newacronym{er}{ER}{Estrogen receptor}
\newacronym{her2}{HER2}{Human epidermal growth factor receptor 2}
\newacronym{svd}{SVD}{Singular value decomposition}
\newacronym{rma}{RMA}{Robust Multi Array}
\newacronym{mas}{MAS5}{Microarray Analysis Suite 5.0}
\newacronym{fdr}{FDR}{False Discovery Rate}
\newacronym{go}{GO}{Gene Ontology}
\newacronym{usa}{USA}{United States of America}
\newacronym{npy}{NPY}{neuropeptide Y}
\newacronym{agrp}{AGRP}{agouti-related protein}
\newacronym{pomc}{POMC}{pro-opiomelanocortin}
\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
\newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
\newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
\newacronym{msh}{MSH}{melanocyte-stimulating hormone}
\newacronym{nyr}{Y1R}{neuropeptide Y1 receptor}
\newacronym{nyr}{Y2R}{neuropeptide Y2 receptor}
\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
\newacronym{irs}{IRS}{insulin receptor seubstrate}
\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
\newacronym{jak}{JAK}{janus kinase}
\newacronym{pc1}{PC1}{prohormone convertase 1}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{snp}{SNP}{single nucleotide polymorphism}
\newacronym{rb}{RB}{retinoblastoma-associated}
\newacronym{opa}{OPA}{obesity policy action}
\newacronym{g1}{G1}{Gap 1}
\newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}
\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
\newacronym{stat}{STAT}{signal transducer and activator of transcription}
\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}

>>>>

\newglossaryentry{ldp}{
	name=Limited duration prevalence,
	text=limited duration prevalence,
	description={Number of patients diagnosed with a disease within a fixed period of time in the past \citep{Bray2013}}
}

@@

\newglossaryentry{chromothripsis}{
	name=Chromothripsis,
	text=chromothripsis,
	description={An all-at-once mutational pattern in which a large number of rearrangements are confined to local regions of one or more chromosomes \citep{Stephens2011}}
}

\newglossaryentry{kataegis}{
	name=Kataegis,
	text=kataegis,
	description={Showers of single nucleotide changes confined to local regions, often near sites of chromosome rearrangements \citep{Stephens2011}}
}

@@

\newacronym{blca}{BLCA}{bladder urothelial carcinoma}
\newacronym{cesc}{CESC}{cervical squamous cell carcinoma and endocervical adenocarcinoma}
\newacronym{coad}{COAD}{colon adenocarcinoma}
\newacronym{kirp}{KIRP}{kidney renal clear cell carcinoma}
\newacronym{lihc}{LIHC}{liver hepatocellular carcinoma}
\newacronym{read}{READ}{rectum adenocarcinoma}
\newacronym{skcm}{SKCM}{skin cutaneous melanoma}
\newacronym{ucec}{UCEC}{uterine corpus endometrial carcinoma}
\newacronym{geo}{GEO}{Gene Expression Omnibus}
\newacronym{tcga}{TCGA}{The Cancer Genome Atlas}
\newacronym{icgc}{ICGC}{International Cancer Genome Consortium}
\newacronym{bmi}{BMI}{Body Mass Index}
\newacronym{who}{WHO}{World Health Organization}
\newacronym{ln}{LN}{Lymph node}
\newacronym{pr}{PR}{Progesterone receptor}
\newacronym{er}{ER}{Estrogen receptor}
\newacronym{her2}{HER2}{Human epidermal growth factor receptor 2}
\newacronym{svd}{SVD}{Singular value decomposition}
\newacronym{rma}{RMA}{Robust Multi Array}
\newacronym{mas}{MAS5}{Microarray Analysis Suite 5.0}
\newacronym{fdr}{FDR}{False Discovery Rate}
\newacronym{go}{GO}{Gene Ontology}
\newacronym{usa}{USA}{United States of America}
\newacronym{npy}{NPY}{neuropeptide Y}
\newacronym{agrp}{AGRP}{agouti-related protein}
\newacronym{pomc}{POMC}{pro-opiomelanocortin}
\newacronym{cart}{CART}{cocaine and amphetamine related transcript}
\newacronym{mc1r}{MC1R}{melanocortin 1 receptor}
\newacronym{mc4r}{MC4R}{melanocortin 4 receptor}
\newacronym{msh}{MSH}{melanocyte-stimulating hormone}
\newacronym{nyr1}{Y1R}{neuropeptide Y1 receptor}
\newacronym{acth}{ACTH}{adrenocorticotropic hormone}
\newacronym{ghsr}{GHSR}{growth hormone secretagogue receptor}
\newacronym{gaba}{GABA}{$\gamma$-aminobutyric acid}
\newacronym{irs}{IRS}{insulin receptor seubstrate}
\newacronym{pi3k}{PI3K}{phosphatidylinositol 3-kinase}
\newacronym{jak}{JAK}{janus kinase}
\newacronym{pc1}{PC1}{prohormone convertase 1}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{fto}{FTO}{fat mass and obesity associated}
\newacronym{snp}{SNP}{single nucleotide polymorphism}
\newacronym{rb}{RB}{retinoblastoma-associated}
\newacronym{opa}{OPA}{obesity policy action}
\newacronym{g1}{G1}{Gap 1}
\newacronym{bcl2}{Bcl-2}{B-cell lymphoma 2}
\newacronym{bh3}{BH3}{Bcl-2 homology 3}
\newacronym{dna}{DNA}{deoxyribonucleic acid}
\newacronym{vegf}{VEGF}{vascular endothelial growth factor}
\newacronym{vegfa}{VEGF-A}{vascular endothelial growth factor-A}
\newacronym{tsp1}{TSP-1}{thrombospondin-1}
\newacronym{emt}{EMT}{endothelial-mesenchymal transition}
\newacronym{met}{MET}{mesenchymal-endothelial transition}
\newacronym{aid}{AID}{activation-induced cytidine deaminase}
\newacronym{apobec}{APOBEC}{apolipoprotein B mRNA editing enzyme, catalytic polypeptide-like}
\newacronym{ssdna}{ssDNA}{single-stranded DNA}
\newacronym{tgfb}{TGF-$\beta$}{transforming growth factor-$\beta$}
\newacronym[longplural={genome wide association studies},\glsshortpluralkey={GWAS}]{gwas}{GWAS}{genome wide association study}
\newacronym{stat}{STAT}{signal transducer and activator of transcription}
\newacronym{pyy}{PYY$_{3-36}$}{peptide YY$_{3-36}$}
\newacronym[longplural={Disability-Adjusted Life Years}]{daly}{DALY}{Disability-Adjusted Life Year}
